---
title: "Penn births cleanup"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

WHYY project data cleanup. Problem was inconsistent data in multiple sheets that needed to be combined into single data frame.

```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(readxl)
library(purrr)
```

## Data problems

These fixes were done in Excel before processing:

- Added consistent header row to each sheet.
- Removed excess header rows.
- Fixed combined cesarean columns
- Removed 1999 and earlier for various data integrity issues

The exported data still needs cleaning on hospital names and possibly county names.

## Import

This comes from [read-xl workflows](https://readxl.tidyverse.org/articles/articles/readxl-workflows.html#concatenate-worksheets-into-one-data-frame-1).

```{r}
path <- "data-raw/births-by-hospital-cleaned.xlsx"
sheets <- excel_sheets(path)
births <-
  lapply(excel_sheets(path), read_excel, path = path)
births <- lapply(seq_along(sheets), function(i) {
  data.frame(sheet = I(sheets[i]), births[[i]])
})
births <- do.call(rbind, births)

births %>% glimpse()
```

## Cleaning

Made county and hospital values uppercase. Pulled year from sheet name.

```{r}
births <- births %>% 
  mutate(
    county = county %>% str_to_upper(),
    hospital = hospital %>% str_to_upper(),
    year = sheet %>% str_sub(1,4)
  ) %>% 
  select(-sheet) %>% 
  select(year, everything())

births %>% glimpse()
```

## Export

```{r}
births %>% write_excel_csv("data-processed/penn-births.csv")

births %>% 
  filter(
    county == "PHILADELPHIA"
  ) %>% 
  write_excel_csv("data-processed/phi-births.csv")
```

